---
title: "Project 1"
author: "YS"
date: "2023-02-23"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Childhood stunting

Childhood stunting is a global public health issue. Stunting is
determined by comparing children's height to a reference population of
children of "normal" nutrition levels.

I found the data set from United Nations Children's Fund (UNICEF). The
provided dataset contains the following variables:

-   WHO Regions (**who_regions**)
-   World Bank Income Group (**wbincgrp**)
-   Prevalence of biologically male children that are stunted
    (**maleprev**)
-   Prevalence of biologically female children that are stunted
    (**femaleprev**)
-   Prevalence of urban children that are stunted (**urbanprev**)
-   Prevalence of rural children that are stunted (**ruralprev**)
-   Prevalence of mothers with no education (**mneduprev**)

I want to evaluate **what numeric predictors** are significantly
associated with the national prevalence (**natprev**).

### Step 1: load the package(s)

```{r}
library(tidyverse)
library(tables)
```

### Step 2: load in the data

```{r data}
df <- read_csv("./Data/UNICEF.csv")
df
```

### Step 3: summary statistics of each numeric variable

1.  national prevalence (natprev)

```{r}
cat("natprev_mean:", mean(df$natprev, trim = 0, na.rm = FALSE))
```

```{r}
cat("natprev_median:", median(df$natprev, trim = 0, na.rm = FALSE))
```

```{r}
natprev_max <- max(df$natprev, na.rm = FALSE)
natprev_min <- min(df$natprev, na.rm = FALSE)
range1 <- (natprev_max - natprev_min)
cat("natprev_range:", range1)
```

```{r}
cat("natprev_sd:", natprev_sd <- sd(df$natprev, na.rm = FALSE))
```

2.  Prevalence of biologically male children that are stunted (maleprev)

```{r}
cat("maleprev_mean:", mean(df$maleprev, trim = 0, na.rm = TRUE))
```

```{r}
cat("maleprev_median:", median(df$maleprev, trim = 0, na.rm = TRUE))
```

```{r}
maleprev_max <- max(df$maleprev, na.rm = TRUE)
maleprev_min <- min(df$maleprev, na.rm = TRUE)
range2 <- (maleprev_max - maleprev_min)
cat("maleprev_range:", range2)
```

```{r}
cat("maleprev_sd:", sd(df$maleprev, na.rm = TRUE))
```

3.  Prevalence of biologically female children that are stunted
    (femaleprev)

```{r}
cat("femaleprev_mean:", mean(df$femaleprev, trim = 0, na.rm = TRUE))
```

```{r}
cat("femaleprev_median:", median(df$femaleprev, trim = 0, na.rm = TRUE))
```

```{r}
femaleprev_max <- max(df$femaleprev, na.rm = TRUE)
femaleprev_min <- min(df$femaleprev, na.rm = TRUE)
range3 <- (femaleprev_max - femaleprev_min)
cat("femaleprev_range:", range3)
```

```{r}
cat("femaleprev_sd:", sd(df$femaleprev, na.rm = TRUE))
```

4. Prevalence of urban children that are stunted (urbanprev)

```{r}
cat("urbanprev_mean:", mean(df$urbanprev, trim = 0, na.rm = TRUE))
```

```{r}
cat("urbanprev_median:", median(df$urbanprev, trim = 0, na.rm = TRUE))
```
```{r}
urbanprev_max <- max(df$urbanprev, na.rm = TRUE)
urbanprev_min <- min(df$urbanprev, na.rm = TRUE)
range4 <- (urbanprev_max - urbanprev_min)
cat("urbanprev_range:", range4)
```

```{r}
cat("urbanprev_sd:", sd(df$urbanprev, na.rm = TRUE))
```

5. Prevalence of rural children that are stunted (ruralprev)







## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to
prevent printing of the R code that generated the plot.
